Approximation as Differentiation: Notes

### 2025-05-08 main/formal-structures.tex

\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
in $\LatGal$, $f$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
$(X, \meet, \top)$. Similarly, $g$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.

@@

For a total language where types denote sets, our goal is to interpret programs as Galois connections relating
approximations of their inputs to approximations of their outputs. Since these lattice of approximations
depends on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
component, assigning $\tau$ to a set $X$, and an associated lattice of approximations $\partial X(x)$ for
every point $x \in X$. The semantics of expression $e$ of type $Y$ with a single free variable of type $X$
must then also have two components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of
the expression, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for
every $x \in X$ which can be used for slicing over that expression.

@@

\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the functor $T(f)$. The composition $(g,
\partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g \comp f,
T(f)(\partial g) \comp \partial f)$.

>>>>

\noindent Right adjoints preserves limits and left adjoints preserves colimits, so for any $f \adj g: X \to Y$
in $\LatGal$, $g$ is a \emph{meet-semilattice homomorphism}, i.e.~preserves the meet-semilattice structure
$(X, \meet, \top)$. Similarly, $f$ is a join-semilattice homomorphism with respect to $(X, \join, \bot)$.

@@

For a total language where types denote sets, our goal is to interpret programs as Galois connections relating
approximations of their inputs to approximations of their outputs. Since these lattices of approximations
depends on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
component, assigning $\tau$ to a set $X$, and an associated lattice of approximations $\partial X(x)$ for
every point $x \in X$. The semantics of expression $e$ of type $Y$ with a single free variable of type $X$
must then also have two components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of
the expression, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for
every $x \in X$ which can be used for slicing over that expression.

@@

\noindent where $T(f)(\partial g)$ is the image of $\partial g$ in the ``reindexing'' functor $T(f)$. The
composition $(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is then given by $(g
\comp f, T(f)(\partial g) \comp \partial f)$.

### 2025-05-08 notes/cmon-enriched.tex

\noindent Now suppose $\mathscr{V}$ is monoidal closed. Then the hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op
\to \mathscr{V}$ is itself $\mathscr{V}$-enriched, providing a family of morphisms $\Hom{\cat{C}}{-}{X}_{Y,Z}:
\Hom{\cat{C}}{Y}{Z} \to \Hom{\mathscr{V}}{\Hom{\cat{C}}{Z}{X}}{\Hom{\cat{C}}{Y}{X}}$ in $\mathscr{V}$ for any
$Y, Z$ in $\cat{C}$. Moreover there exists an enriched Yoneda embedding, namely a functor
$\Yoneda_{\mathscr{V}}: \cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ which is itself
$\mathscr{V}$-enriched, sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$
and providing a family of morphisms $(\Yoneda_{\mathscr{V}})_{X,Y}: \Hom{\cat{C}}{X}{Y} \to $ to
$\mathscr{V}$-natural transformations $\Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each
component as post-composition with $f$.

>>>>

\noindent Now suppose $\mathscr{V}$ is monoidal closed. Then the hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op
\to \mathscr{V}$ is itself $\mathscr{V}$-enriched, providing a family of morphisms $\Hom{\cat{C}}{-}{X}_{Y,Z}:
\Hom{\cat{C}}{Y}{Z} \to \Hom{\mathscr{V}}{\Hom{\cat{C}}{Z}{X}}{\Hom{\cat{C}}{Y}{X}}$ in $\mathscr{V}$ for any
$Y, Z$ in $\cat{C}$. Moreover there exists an enriched Yoneda embedding, namely a functor
$\Yoneda_{\mathscr{V}}: \cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ which is itself
$\mathscr{V}$-enriched, sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$
and providing a family of morphisms $(\Yoneda_{\mathscr{V}})_{X,Y}: \Hom{\cat{C}}{X}{Y} \to
\Hom{\Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}}{\Hom{\cat{C}}{-}{X}}{\Hom{\cat{C}}{-}{Y}}$ acting
concretely at each component as post-composition with $f: X \to Y$.
